---
title: "Historical dataset comparison"
author: "Amber Runyon"
date: "10/18/2021"
output: html_document
---
This script compares distributions between MACA historical and gridmet to see if we can use observed historical for comparison instead of projected to significantly speed up download times.

## First pull data

```{r, echo=FALSE, message=FALSE}
rm(list=ls())
library(dplyr)
library(ggplot2)
data.dir <- "C:/Users/achildress/DOI/NPS-NRSS-CCRP-FC Science Adaptation - Documents/General/RSS Stuff/Parks/CARE/RCF with WB1.1_WW HD/data/park-specific/"
load(paste0(data.dir,"output/CARE_38.383553_-111.283004_Final_Environment.RData"))

grid <- read.csv(paste0(data.dir,"/input/GridMet.csv"))
```

MACA projected historical period = 1950-2005

Gridmet period is 1979-present, but period used for training MACA = 1979-2012 
https://climate.northwestknowledge.net/MACA/MACAtrainingdata.php

According to https://climate.northwestknowledge.net/MACA/MACAfaq.php, at least 30 years of observed data should be used, it it's preferable to use whole reference to avoid statistical differences of change.

```{r, echo=FALSE, message=FALSE}
grid %>% mutate(Date = as.Date(grid$Date,format="%Y-%m-%d"),
                yr = format(grid$Date,format="%Y")) %>%
  filter(yr < 2012) -> grid
  

ALL_HIST %>% mutate(Date = as.Date(ALL_HIST$Date,format="%Y-%m-%d"),
                yr = format(ALL_HIST$Date,format="%Y")) %>%
  select(Date,GCM,PrecipCustom,TmaxCustom,TminCustom,yr) -> AH

aggregate(cbind(PrecipCustom,TmaxCustom)~GCM,AH,mean)
c<- aggregate(cbind(TmaxCustom,TminCustom,PrecipCustom)~GCM,AH,summary)

AH %>% group_by(GCM) %>%
    summarise(across(starts_with('val'), mean))
summary(grid)

ggplot(data=AH, aes(x=PrecipCustom, group=GCM, fill=GCM)) +
    geom_density(adjust=1.5) +
    # theme_ipsum() +
    facet_wrap(~GCM) +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      axis.ticks.x=element_blank()
    )


```